### 2013-05-04 writing/paper/abstract.tex
reformatted thesis, incorporating lots of oli's edits

\begin{abstractpage}
Intraplate faulting in central Asia is a major component of the Indo-Eurasian collision. The kinematics and mechanisms of intraplate deformation are important in understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources. I examine the fault kinematics surrouding the 150km wide Shargyn Basin at the intersection of the left-lateral transpressional Gobi-Altai fault system and the right-lateral transpressional Altai fault system. The basin is formed by a compressional stepover, an uplifted wedge from the intersecting strike-slip systems and many terminal thrust splays. The fault-kinematics, pre-existing structural weaknesses and a significant gravity anomaly all suggest that the basin basement is mechanically strong. This research demonstrates some of the potential kinematics for intraplate transpressional orogenies and emphasizes the importance of pre-existing crustal structure in the development of active faults.
\end{abstractpage}