### 2013-05-04 writing/paper/main.tex

\documentclass[12pt, doublepage, vi]{mitthesis}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[demo]{graphicx}
\usepackage{natbib}
\usepackage[margin=0.75in]{geometry}
\usepackage{setspace}
\renewcommand\thechapter{}
\renewcommand\thesection{\arabic{section}}
\renewcommand\thesubsection{\thesection.\arabic{subsection}}
\begin{document}
\title{Active deformation of the Shargyn Basin, a transpressional strike-slip intersection in western Mongolia.}
\author{Thomas Ben Thompson}
\department{Department of Earth, Atmospheric and Planetary Science}
\degree{Bachelor of Science in Earth, Atmospheric and Planetary Sciences}
\degreemonth{June}
\degreeyear{2013}
\copyrightnoticetext{\copyright~ 2013 Thomas Ben Thompson \newline
Anyone is permitted to reproduce and distribute copies of thesis so long as they give me credit.}
\thesisdate{May 10, 2013}
\supervisor{Oliver Jagoutz}{Assistant Professor}
\chairman{Samuel Bowring}{Chair, Committee on Undergraduate Program}
\maketitle
\include{abstract}

\section*{Acknowledgments}
Primarily, I thank Claire Bucholz (EAPS PhD candidate) for wandering the Mongolian steppes and mountains and teaching me endlessly about geology and EAPS Professor Oliver Jagoutz for supervising and helping me through the whole thesis process. I thank the UROP office and the EAPS department for supporting my field work in Mongolia last summer.

\tableofcontents
\listoffigures
\clearpage
\include{intro}
\include{background}
\include{structure}
\include{geophysical}
\include{basin}
\include{conclusion}
\bibliographystyle{plainnat}
\bibliography{/home/tbent/projects/library.bib}

\end{document}