### 2013-10-07 writing/jaes_article/main.tex

\documentclass[12pt]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[margin=1.0in]{geometry}
\usepackage{setspace}
\usepackage{caption}
\doublespacing
\AtBeginCaption{\doublespacing}
\begin{document}
\title{Active deformation of the Shargyn Basin, a transpressional strike-slip intersection in western Mongolia.}
\author{%
    Thomas Ben Thompson (Corresponding Author)\\
    Department of Earth, Atmospheric and Planetary Sciences at MIT\\
    \texttt{tthompson@fas.harvard.edu}\vspace{40pt} \\
    Oliver Jagoutz \\
    Department of Earth, Atmospheric and Planetary Sciences at MIT \\
    \texttt{jagoutz@mit.edu} \vspace{40pt} \\
    Claire Bucholz \\ 
    Department of Earth, Atmospheric and Planetary Sciences at MIT \\
    \texttt{cbucholz@mit.edu} \\
    }
\maketitle

\include{abstract}

\tableofcontents
\listoffigures
\clearpage
\include{intro}
\include{background}
\include{structure}
\include{discussion}
\include{conclusion}
\appendix
\include{xsecappendix}
\section*{Acknowledgments}
Many thanks to Yeson Erdene, a student at MUST, and Agi, our friend and driver. 
\bibliographystyle{plainnat}
\bibliography{/home/tbent/projects/library.bib}

\end{document}